---
title: "SDM4 in R: Multiple Regression (Chapter 28)"
author: "Nicholas Horton (nhorton@amherst.edu)"
date: "February 6, 2016"
output: 
  pdf_document:
    fig_height: 3
    fig_width: 6
  html_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 4
    fig_width: 6
---


```{r, include=FALSE}
# Don't delete this chunk if you are using the mosaic package
# This loads the mosaic and dplyr packages
require(mosaic)
```

```{r, include=FALSE}
# Some customization.  You can alter or delete as desired (if you know what you are doing).

# This changes the default colors in lattice plots.
trellis.par.set(theme=theme.mosaic())  

# knitr settings to control how R chunks work.
require(knitr)
opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small"    # slightly smaller font for code
)
```

## Introduction and background 

This document is intended to help describe how to undertake analyses introduced 
as examples in the Fourth Edition of \emph{Stats: Data and Models} (2014) by De Veaux, Velleman, and Bock.
More information about the book can be found at http://wps.aw.com/aw_deveaux_stats_series.  This
file as well as the associated R Markdown reproducible analysis source file used to create it can be found at http://www.amherst.edu/~nhorton/sdm4.

This work leverages initiatives undertaken by Project MOSAIC (http://www.mosaic-web.org), an NSF-funded effort to improve the teaching of statistics, calculus, science and computing in the undergraduate curriculum. In particular, we utilize the `mosaic` package, which was written to simplify the use of R for introductory statistics courses. A short summary of the R needed to teach introductory statistics can be found in the mosaic package vignettes (http://cran.r-project.org/web/packages/mosaic).

## Chapter 28: Multiple Regression

### Section 28.1: What is multiple regression?


```{r}
library(mosaic); library(readr)
options(na.rm=TRUE)
options(digits=3)
(6.54 - 5.91)/0.56  # should be 1.1 sd better, see page 112
```

```{r}
Heptathlon <- 
read_delim("http://www.amherst.edu/~nhorton/sdm4/data/Womens_Heptathlon_2012.txt",
  delim="\t")
nrow(Heptathlon)  
filter(Heptathlon, LJ >= max(LJ, na.rm=TRUE)) %>% data.frame()
favstats(~ LJ, data=Heptathlon)
(6.54 - mean(~ LJ, data=Heptathlon))/sd(~ LJ, data=Heptathlon)
```

### Section 28.2: Interpreting multiple regression coefficients

### Section 28.3: The multiple regression model (assumptions and conditions)

### Section 28.4: Multiple regression inference

### Section 28.5: Comparing multiple regression models

