---
title: "SDM4 in R: Multiple Regression Wisdom (Chapter 29)"
author: "Nicholas Horton (nhorton@amherst.edu)"
date: "April 3, 2016"
output: 
  pdf_document:
    fig_height: 2.8
    fig_width: 7
  html_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 4
    fig_width: 6
---


```{r, include=FALSE}
# Don't delete this chunk if you are using the mosaic package
# This loads the mosaic and dplyr packages
require(mosaic)
options(digits=3)
```

```{r, include=FALSE}
# Some customization.  You can alter or delete as desired (if you know what you are doing).

# This changes the default colors in lattice plots.
trellis.par.set(theme=theme.mosaic())  

# knitr settings to control how R chunks work.
require(knitr)
opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small"    # slightly smaller font for code
)
```

## Introduction and background 

This document is intended to help describe how to undertake analyses introduced 
as examples in the Fourth Edition of \emph{Stats: Data and Models} (2014) by De Veaux, Velleman, and Bock.
More information about the book can be found at http://wps.aw.com/aw_deveaux_stats_series.  This
file as well as the associated R Markdown reproducible analysis source file used to create it can be found at http://www.amherst.edu/~nhorton/sdm4.

This work leverages initiatives undertaken by Project MOSAIC (http://www.mosaic-web.org), an NSF-funded effort to improve the teaching of statistics, calculus, science and computing in the undergraduate curriculum. In particular, we utilize the `mosaic` package, which was written to simplify the use of R for introductory statistics courses. A short summary of the R needed to teach introductory statistics can be found in the mosaic package vignettes (http://cran.r-project.org/web/packages/mosaic).

## Chapter 29: Multiple Regression Wisdom


```{r}
library(mosaic); library(readr)
Coasters <- read_csv("http://www.amherst.edu/~nhorton/sdm4/data/Roller_coasters_2014.csv")
# Name,Park,Track,Speed,Height,Drop,Length,Duration,Inversions,Opened
glimpse(Coasters)
tally(~ Speed, data=Coasters)
filter(Coasters, Speed=="*100" | Speed=="*56")
```
Note that two of the observations for speed have special characters in them.
We first need to address these data irregularities:
```{r}
library(tidyr)
Coasters <- mutate(Coasters, Speed = extract_numeric(Speed))
```
To match the output we need to remove these two coasters.

```{r}
dim(Coasters)
Coasters <- filter(Coasters, 
  !(Name %in% c("Tower of Terror", "Wild Beast")))
dim(Coasters)
```


Figure 29.1 (page 860) displays the scatterplot of duration as a function
of speed.

```{r}
xyplot(Duration ~ Speed, type=c("p", "r"), xlab="Speed (mph)",
  ylab="Duration (sec)", data=Coasters)
```

We can also replicate the regression model (and regression diagnostics)
on the bottom of page 860.

```{r}
mod1 <- lm(Duration ~ Speed, data=Coasters)
msummary(mod1)
xyplot(resid(mod1) ~ fitted(mod1), xlab="Predicted (sec)",
  ylab="Residuals (sec)", type=c("p", "r"))
```

### Section 29.1: Indicators

The model displayed on the bottom of page 861 is restricted to 
roller coasters with no inversions.
```{r}
noinversion <- filter(Coasters, Inversions==0)  # 0 is No (or FALSE)
mod2 <- lm(Duration ~ Speed, data=noinversion)
msummary(mod2)
```

We can do the same for coasters with inversions:
```{r}
mod3 <- lm(Duration ~ Speed, data=filter(Coasters, Inversions==1))
msummary(mod3)
```

The model at the top of page 862 adds the Inversion indicator to the
model.

```{r}
mod4 <- lm(Duration ~ Speed + Inversions, data=Coasters)
msummary(mod4)
mod4fun <- makeFun(mod4)
```

We can generate predicted values from this model.
```{r}
mod4fun(Speed=55, Inversions==0)  # Hayabusa
mod4fun(Speed=60.3, Inversions==1)  # Hangman
```

```{r}





