---
title: "SDM4 in R: Inferences for Regression (Chapter 25)"
author: "Nicholas Horton (nhorton@amherst.edu)"
date: "March 28, 2016"
output: 
  pdf_document:
    fig_height: 2.8
    fig_width: 7
  html_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 4
    fig_width: 6
---


```{r, include=FALSE}
# Don't delete this chunk if you are using the mosaic package
# This loads the mosaic and dplyr packages
require(mosaic)
options(digits=3)
```

```{r, include=FALSE}
# Some customization.  You can alter or delete as desired (if you know what you are doing).

# This changes the default colors in lattice plots.
trellis.par.set(theme=theme.mosaic())  

# knitr settings to control how R chunks work.
require(knitr)
opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small"    # slightly smaller font for code
)
```

## Introduction and background 

This document is intended to help describe how to undertake analyses introduced 
as examples in the Fourth Edition of \emph{Stats: Data and Models} (2014) by De Veaux, Velleman, and Bock.
More information about the book can be found at http://wps.aw.com/aw_deveaux_stats_series.  This
file as well as the associated R Markdown reproducible analysis source file used to create it can be found at http://www.amherst.edu/~nhorton/sdm4.

This work leverages initiatives undertaken by Project MOSAIC (http://www.mosaic-web.org), an NSF-funded effort to improve the teaching of statistics, calculus, science and computing in the undergraduate curriculum. In particular, we utilize the `mosaic` package, which was written to simplify the use of R for introductory statistics courses. A short summary of the R needed to teach introductory statistics can be found in the mosaic package vignettes (http://cran.r-project.org/web/packages/mosaic).

## Chapter 25: Inferences for Regression

### Section 25.1: The population and the sample

```{r}
library(mosaic); library(readr)
BodyFat <- read_csv("http://www.amherst.edu/~nhorton/sdm4/data/Body_fat_complete.csv")
dim(BodyFat)
glimpse(BodyFat)
```

We can confirm the coefficients from the model on page 690.
```{r}
BodyFatmod <- lm(PctBF ~ waist, data=BodyFat)
coef(BodyFatmod)
```

### Section 25.2: Assumptions and conditions

We can regenerate the output and figures for the example on pages 692-696.

```{r}
msummary(BodyFatmod)
rsquared(BodyFatmod)
confint(BodyFatmod)    # see page 700
```

```{r}
# Figure 25.4 
xyplot(PctBF ~ waist, xlab="Waist (in.)", 
       type=c("p", "r", "smooth"), data=BodyFat)
# Figure 25.5
xyplot(resid(BodyFatmod) ~ waist, xlab="Waist (in.)", 
       type=c("p", "r", "smooth"), data=BodyFat)
# equiv of Figure 25.6   note that Figure 25.6 refers to the diamonds dataset
xyplot(resid(BodyFatmod) ~ fitted(BodyFatmod), xlab="Predicted values", 
       ylab="Residuals",
       type=c("p", "r", "smooth"), data=BodyFat)
# Figure on bottom of page 695
qqmath(~ resid(BodyFatmod))
```

#### Section 25.6: Confidence intervals for predicted values

We can reproduce Figure 25.12 (page 707) using the `panel.lmbands()` function.

```{r}
xyplot(PctBF ~ waist, xlab="Waist (in.)", 
       panel=panel.lmbands, cex=0.2, data=BodyFat)
```

```{r}
Craters <- read.csv("http://www.amherst.edu/~nhorton/sdm4/data/Craters.csv")
dim(Craters)
Craters <- mutate(Craters,
                  logDiam = log(Diam.km.),
                  logAge = log(age..Ma.))
Cratermod <- lm(logDiam ~ logAge, data=Craters)
favstats(~ logAge, data=Craters)   # note example in book has n=39
```

```{r}
confpred <- predict(Cratermod, interval="confidence")
intpred <- predict(Cratermod, interval="prediction")
select(Craters, -Name) %>% head(., 3)
head(confpred, 3)
head(intpred, 3)
```

#### Section 25.7: Logistic regression

The Pima Indian dataset example is given on pages 708-712.
```{r}
Diabetes <- read_csv("http://www.amherst.edu/~nhorton/sdm4/data/Pima_Indians_Diabetes.csv")
Diabetes <- filter(Diabetes, BMI>0)  # get rid of missing values for BMI
xyplot(Diabetes ~ BMI, type=c("p", "r", "smooth"), data=Diabetes)
bwplot(BMI ~ as.factor(Diabetes), data=Diabetes)
```

```{r}
pimamod <- glm(Diabetes ~ BMI, family="binomial", data=Diabetes)
msummary(pimamod)
```
