---
title: "SDM4 in R: Multifactor Analysis of Variance (Chapter 27)"
author: "Patrick Frenett (pfrenett19@amherst.edu)"
date: "June 8, 2016"
output: 
  pdf_document:
    fig_height: 2.8
    fig_width: 7
  html_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 4
    fig_width: 6
---


```{r, include=FALSE}
# Don't delete this chunk if you are using the mosaic package
# This loads the mosaic and dplyr packages
require(mosaic)
options(digits=5)
```

```{r, include=FALSE}
# Some customization.  You can alter or delete as desired (if you know what you are doing).

# This changes the default colors in lattice plots.
trellis.par.set(theme=theme.mosaic())  

# knitr settings to control how R chunks work.
require(knitr)
opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small"    # slightly smaller font for code
)
```

## Introduction and background 

This document is intended to help describe how to undertake analyses introduced 
as examples in the Fourth Edition of \emph{Stats: Data and Models} (2014) by De Veaux, Velleman, and Bock.
More information about the book can be found at http://wps.aw.com/aw_deveaux_stats_series.  This
file as well as the associated R Markdown reproducible analysis source file used to create it can be found at http://www.amherst.edu/~nhorton/sdm4.

This work leverages initiatives undertaken by Project MOSAIC (http://www.mosaic-web.org), an NSF-funded effort to improve the teaching of statistics, calculus, science and computing in the undergraduate curriculum. In particular, we utilize the `mosaic` package, which was written to simplify the use of R for introductory statistics courses. A short summary of the R needed to teach introductory statistics can be found in the mosaic package vignettes (http://cran.r-project.org/web/packages/mosaic).

## Chapter 27: Multifactor Analysis of Variance

Figure 27.1 on page 782 gives two boxplots.

```{r}
Darts <- read.csv("http://www.amherst.edu/~nhorton/SDM3/Chapter_29/Ch29_Darts.csv")
bwplot(Accuracy~Distance,data=Darts)
bwplot(Accuracy~Hand,data=Darts)
```

### Section 27.1: A Two-Factor ANOVA Model

The summary of the table on page 784

```{r}
lmDarts <- lm(Accuracy ~ Distance + Hand, data=Darts)
summary(aov(lmDarts))
```

### Section 27.2: Assumptions and Conditions

Figures 27.2 and 27.3 showing hand and distance against accuracy (as well as adjusted hand and adjusted distance)

```{r}
lmHand <- lm(Accuracy ~ Hand, data=Darts)
lmDistance <- lm(Accuracy ~ Distance, data=Darts)

bwplot(Accuracy~Hand,data=Darts)
bwplot(resid(lmDistance)~Hand,data=Darts)

bwplot(Accuracy~Distance,data=Darts)
bwplot(resid(lmHand)~Distance,data=Darts)

with(Darts,interaction.plot(Distance,Hand, Accuracy))
```

### Section 27.3: Interactions

Interaction plot and ANOVA summary on page 797.

```{r}
with(Darts,interaction.plot(Hand,Distance, Accuracy))
lmDarts <- lm(Accuracy ~ Distance*Hand, data=Darts)
summary(aov(lmDarts))
```