---
title: "SDM4 in R: Multiple Regression (Chapter 28)"
author: "Nicholas Horton (nhorton@amherst.edu)"
date: "January 2, 2017"
output: 
  pdf_document:
    fig_height: 3
    fig_width: 6
  html_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 4
    fig_width: 6
---


```{r, include=FALSE}
# Don't delete this chunk if you are using the mosaic package
# This loads the mosaic and dplyr packages
require(mosaic)
```

```{r, include=FALSE}
# Some customization.  You can alter or delete as desired (if you know what you are doing).

# This changes the default colors in lattice plots.
trellis.par.set(theme=theme.mosaic())  

# knitr settings to control how R chunks work.
require(knitr)
opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small"    # slightly smaller font for code
)
```

## Introduction and background 

This document is intended to help describe how to undertake analyses introduced 
as examples in the Fourth Edition of \emph{Stats: Data and Models} (2014) by De Veaux, Velleman, and Bock.
More information about the book can be found at http://wps.aw.com/aw_deveaux_stats_series.  This
file as well as the associated R Markdown reproducible analysis source file used to create it can be found at http://nhorton.people.amherst.edu/sdm4.

This work leverages initiatives undertaken by Project MOSAIC (http://www.mosaic-web.org), an NSF-funded effort to improve the teaching of statistics, calculus, science and computing in the undergraduate curriculum. In particular, we utilize the `mosaic` package, which was written to simplify the use of R for introductory statistics courses. A short summary of the R needed to teach introductory statistics can be found in the mosaic package vignettes (http://cran.r-project.org/web/packages/mosaic).

## Chapter 28: Multiple Regression

### Section 28.1: What is multiple regression?


The table on page 818 displays the results from the multiple regression model.

```{r message=FALSE}
library(mosaic); library(readr)
options(digits=3)
BodyFat <- read_csv("http://nhorton.people.amherst.edu/sdm4/data/Body_fat_complete.csv")
BodyFatmod <- lm(PctBF ~ waist + Height, data=BodyFat)
msummary(BodyFatmod)
```

### Section 28.2: Interpreting multiple regression coefficients

Figure 28.1 on page 819 displays the scatterplot of percent body fat against height.

```{r}
xyplot(PctBF ~ Height, type=c("p", "r"), data=BodyFat)
```

Figure 28.2 (page 820) displays the scatterplot for a subset of the data (men with waist sizes between 36 and 38 inches).

```{r}
xyplot(PctBF ~ Height, type=c("p", "r"), data=filter(BodyFat, waist > 36, waist < 38))
```

Figure 28.3 (page 820) displays the partial regression plot for weight.

```{r}
BodyFatwaist <- lm(PctBF ~ waist, data=BodyFat)
BodyFatheight <- lm(Height ~ waist, data=BodyFat)

xyplot(resid(BodyFatwaist) ~ resid(BodyFatheight), 
       ylab="% body fat residuals", xlab="Height residuals", type=c("p", "r"))
```

### Section 28.3: The multiple regression model (assumptions and conditions)

Figure 28.4 (page 822) displays scatterplots of residuals vs. height and waist, respectively.

```{r}
xyplot(resid(BodyFatmod) ~ Height, type=c("p", "r"), data=BodyFat)
xyplot(resid(BodyFatmod) ~ waist, type=c("p", "r"), data=BodyFat)
```

Figure 28.5 (page 823) displays histogram and qq plot of the residuals.

```{r}
histogram(~ resid(BodyFatmod), fit="normal")
qqmath(~ resid(BodyFatmod))
```

 
### Section 28.4: Multiple regression inference

Figure 28.6 (page 829) displays the scatterplot matrix infant mortality data.  

```{r, fig.height=6,message=FALSE}
InfantMortality <- read_csv("http://nhorton.people.amherst.edu/sdm4/data/Infant_Mortality.csv")
splom(select(InfantMortality, -State))
```

In addition, we display a scatterplot matrix for the motivating example from the chapter (BodyFat) using the `GGally` package.
```{r, message=FALSE, fig.height=6}
subsetBodyFat <- select(BodyFat, PctBF, Height, waist)
library(GGally)
ggpairs(subsetBodyFat)
```

 
### Section 28.5: Comparing multiple regression models

We may want to compare which of our models provides the most parsimonious fit to these data.

```{r}
msummary(BodyFatheight)
msummary(BodyFatwaist)
msummary(BodyFatmod)
```

The adjusted R-squared value of 0.711 is considerably higher for the model
with both predictors (though the model with just waist has an adjusted R-squared value of 0.677).
