---
title: "SDM4 in R: Regression Wisdom (Chapter 8)"
author: "Nicholas Horton (nhorton@amherst.edu)"
date: "June 4, 2018"
output: 
  pdf_document:
    fig_height: 3
    fig_width: 6
  html_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 4
    fig_width: 6
---


```{r, include = FALSE}
# Don't delete this chunk if you are using the mosaic package
# This loads the mosaic and dplyr packages
require(mosaic)
```

```{r, include = FALSE}
# knitr settings to control how R chunks work.
require(knitr)
opts_chunk$set(
  tidy = FALSE,     # display code as typed
  size = "small"    # slightly smaller font for code
)
```

## Introduction and background 

This document is intended to help describe how to undertake analyses introduced 
as examples in the Fourth Edition of \emph{Stats: Data and Models} (2014) by De Veaux, Velleman, and Bock.
More information about the book can be found at http://wps.aw.com/aw_deveaux_stats_series.  This
file as well as the associated R Markdown reproducible analysis source file used to create it can be found at http://nhorton.people.amherst.edu/sdm4.

This work leverages initiatives undertaken by Project MOSAIC (http://www.mosaic-web.org), an NSF-funded effort to improve the teaching of statistics, calculus, science and computing in the undergraduate curriculum. In particular, we utilize the `mosaic` package, which was written to simplify the use of R for introductory statistics courses. A short summary of the R needed to teach introductory statistics can be found in the mosaic package vignettes (http://cran.r-project.org/web/packages/mosaic).
A paper describing the mosaic approach was published in the *R Journal*: https://journal.r-project.org/archive/2017/RJ-2017-024.


## Chapter 8: Regression Wisdom

### Section 8.1: Examining residuals


Figure 8.1 (page 220) displays the scatterplot of heart rate vs duration for the Penguins dataset (along with a superimposed regression line and a smoother).

```{r message = FALSE}
library(mosaic)
library(readr)
options(digits = 3)
Penguins <- read_csv("http://nhorton.people.amherst.edu/sdm4/data/Penguins.csv")
gf_point(DiveHeartRate ~ Duration, ylab = "Dive Heart Rate (bpm)", 
         xlab = "Duration (mins)", data = Penguins) %>%
  gf_lm() %>%
  gf_smooth(se = FALSE)
```

Figure 8.2 (page 220) displays the residuals from a linear regression model as a function of duration.

```{r message=FALSE}
Penguinmod <- lm(DiveHeartRate ~ Duration, data = Penguins)
gf_point(resid(Penguinmod) ~ Duration, data = Penguins) %>%
  gf_lm() %>%
  gf_smooth(se = FALSE)
```

Figure 8.3 (page 221) displays the histogram of residuals for the cereal data from Chapter 7.

```{r message=FALSE}
Cereals <- read_csv("http://nhorton.people.amherst.edu/sdm4/data/Cereals.csv")
Cerealmod <- lm(calories ~ sugars, data = Cereals)
gf_histogram(~ resid(Cerealmod), binwidth = 7.5)
```

Figure 8.4 (page 221) displays a scatterplot of the residuals vs predicted values.  Without jittering, the display has an odd pattern.

```{r}
gf_point(resid(Cerealmod) ~ fitted(Cerealmod))
```

By adding some random noise we can more easily observe values that are shared by more than one cereal.

```{r}
gf_point(jitter(resid(Cerealmod)) ~ jitter(fitted(Cerealmod)))
```

Figure 8.5 (page 222) displays the scatterplot stratified by what shelf it is displayed on at the store.

```{r}
tally(~ shelf, data = Cereals)
Cereals <- mutate(Cereals, shelfgrp = 
  derivedFactor(
    bottomshelf = shelf == 1,
    middleshelf = shelf == 2,
    topshelf = shelf == 3
  )
)
tally(~ shelfgrp, data = Cereals)
gf_point(calories ~ sugars, color = ~ shelfgrp, 
  lwd = 2, data = Cereals) %>%
  gf_lm()
```


### Section 8.2: Extrapolation and reaching beyond the data

### Section 8.3: Outliers, leverage, and influence

### Section 8.4: Lurking variables and causation

### Section 8.5: Working with summary values

